---
title: "utaustin data"
author: "Pranaya Jajoo"
date: "2023-06-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

pulling data
```{r}
download.file("http://www.openintro.org/stat/data/evals.RData", destfile = "evals.RData")
load("evals.RData")
head(evals)
summary(evals)
```

looking at the distribution of the score
```{r}
hist(evals$score)
```

looking at possible correlation between beauty and score
```{r}
score_bty_lm = lm(evals$score ~ evals$bty_avg)
summary(score_bty_lm)
plot(jitter(evals$score,factor=1.2) ~ jitter(evals$bty_avg,factor=1.2))
abline(score_bty_lm)
cor(evals$score, evals$bty_avg)
```

evaluating the conditions of least squares regression
```{r}
# variablity - constant
plot(score_bty_lm$residuals ~ evals$bty_avg)
abline(h = 0, lty = 4)  # adds a horizontal dashed line at y = 0
# normality - left skewed
hist(score_bty_lm$residuals)
# linearity - data is positively linear
qqnorm(score_bty_lm$residuals)
qqline(score_bty_lm$residuals)
```

multiple linear regression using beauty and gender
```{r}
m_bty_gen <- lm(score ~ bty_avg + gender, data = evals)
```

evaluating the conditions for regression
```{r}
# normality -- qq plot
qqnorm(m_bty_gen$residuals)
qqline(m_bty_gen$residuals)
# variablity -- residual plot
plot(m_bty_gen$residuals ~ evals$bty_avg)
abline(h = 0, lty = 4)
# box plot
plot(m_bty_gen$residuals ~ evals$gender)
abline(h = 0, lty = 4)  # adds a horizontal dashed line at y = 0
# historgram
hist(m_bty_gen$residuals)

plot(m_bty_gen)

# linearity
plot(jitter(evals$score) ~ evals$bty_avg)
# box plot
plot(evals$score ~ evals$gender)

# residuals are left skewed
# residuals have constant variability
```

formula for score with this multiple regression model is
score = 3.74734 + 0.07416 * bty_avg  + 0.17239 * gendermale
where gendermale = 1 for male
thus, for the same beauty rating, male professors will gave higher score

# testing linear model with beauty and rank
```{r}
lm_bty_rank = lm(score ~ bty_avg + rank, data = evals)
summary(lm_bty_rank)
```
for rank, the model is considering the 3 levels as two different variables

modeling out for all variables to determine which ones are significant and which are insignificant
```{r}
lm_all = lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_profs + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(lm_all)
```
p-value for clas_profs is highest (0.77806), making it the least impactful 


testing by removing clas_profs, the variable with the highest p value
```{r}
lm_no_clsprofs <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval + 
    cls_students + cls_level + cls_credits + bty_avg + pic_outfit + pic_color, 
    data = evals)
summary(lm_no_clsprofs)
`# yes, there was some change. they are now more significant compared to with number of professors variable
```

removing all igh p-value variables and testing the model again to create a final model
```{r}
lm_final <- lm(score ~ gender + language + age + cls_perc_eval + 
    cls_credits + bty_avg + pic_color, data = evals)
summary(lm_final)
```
score = 3.967255 + 0.221457 * gender (where male = 1) - 0.281933 * language (where non-english = 1) - 0.005877 * age + 0.004295 * cls_perc_eval + 0.444392 * cls_credits (where one credit = 1) + 0.048679 * bty_avg - 0.216556 * pic_color (where color = 1)

testing the final models efficacy
```{r}
# Normal
qqnorm(lm_final$residuals)
qqline(lm_final$residuals)

# residuals vs fitted
plot(lm_final)

# histogram
hist(lm_final$residuals)

# linearity
plot(jitter(evals$score) ~ evals$bty_avg)
# boxplot
plot(jitter(evals$score) ~ evals$gender)
plot(jitter(evals$score) ~ evals$ethnicity)
plot(jitter(evals$score) ~ evals$language
     )

plot(jitter(evals$score) ~ evals$age)
plot(jitter(evals$score) ~ evals$cls_perc_eval)
plot(jitter(evals$score) ~ evals$cls_credits)
plot(jitter(evals$score) ~ evals$pic_color)
```

based on the model, to get a high score, the professor should
1. be male
2. have received education in english
3. be younger
4. have more student complete the evaluation
5. should teach subjects with one credit
6. have high beauty rating
7. have a non-color picture







